---
journal: jacsat
type: article
author:
  - name: Rahmi Chowdhury
    aff: Department of Chemistry and Biology, Toronto Metropolitan University, 
    email: \email{rchowdhury@ryerson.ca}
title:
  formatted: Analytical Insights into Thirteen OPE Variables 
  short: An \textsf{achemso} demo
abbr: IR,NMR,UV
keywords: American Chemical Society, \LaTeX
abstract: |
  There are many different parameters that play an integral roll in predictive Organophosphate Esters (OPEs) modelling. Much of this research has revolved around using various physical and chemical properties to determine the Overall Persistance(POV) and Long-Range Transport Potential (LRTP) of these organic chemicals at a screening. While these models have been incredibly beneficial, I wanted to look for ways to better help the predictive patterning of OPEs.To do this, some questions this paper looked to answer was 1) is there a correlation between any of the variables present? 2) How can these relationships (if any) help to improve predictability of OPEs. With that being said, this paper looked to create a basis point for these predictive tools to better understand the relationships between all these parameters. This paper analyzed all the current known available data about each of these parameters for 10 known native OPEs (TPP,TMTP,TPTP,TEP,TPrP,TBP,TBEP,EHDP,TCEP,TCPP) and compared them to one another to see if any sort of relationship can be found. The 14 parameters looked at include: mass, Characteristic Travel Distance in air (CTD_a), Characteristic Travel Distance in water (CTD_w), POV_air, POV_water, Number of Carbon units (C), Number of Hydrogen units(H),  Number of Chlorine units(Cl),  Number of Oxygen units(O), number of Phosphorus(Ph), air-water coefficient from the literature(KAW), octanol-water coefficient from the literature(KOW), and octanol-air coefficient from the literature(KOA), Retention Time	(RT). Using the pearson test, the most statistically signiifcant results include the: POV_w-CTD_w pair,CTD_a-KOA,H-KOA , KAW-POV_a at higher masses, KOW-RT at lower masses,and KOA-mass, they each had values of 0.999,-0.944,0.952,-0.997,0.952, and 0.910 respectively. These relationships are incredibly important because it aids in giving valuable inisght when it comes to predicting the behavior of similar OPEs. Some noteworthy relationships that may come in aid is the strong correlation between KOA values of OPEs and their respective masses. This relationship can help help predict the extent of the LRTP of OPEs. 

bibliography: acstest.bib
csl: american-chemical-society.csl
output: rticles::acs_article



---

\begin{tocentry}
Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 12\,cm by 5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.
\end{tocentry}

# Chapter 1: Introduction  The History of OPEs
Organophosphate Esters (OPEs) have been found all across the world. In food webs through bioaccumulation(Du et al., 2019),  in sediments(Chokwe et al., 2020), in people via absorption, in the air and in the waters up north in the Arctic through various modes of transport (Castro-Jiménez et al., 2014) (Möller et al., 2011) (Sühring et al., 2016)(Salamova et al., 2013) (Salamova et al., 2013) (Luo et al., 2016) (Lai et al., 2015). Recent research suggests these OPEs have persistent organic pollutant (POPs)-like characteristics (Guigueno & Fernie, 2017). While on the human side, studies have shown that some OPEs can potentially be: mutagenic, carcinogenic, genotoxic and neurotoxic (Zhao et al., 2020).  OPEs have the potential to be harmful, also ever since PBDEs were banned, the consumption of these dangerous chemicals has greatly increased (Chupeau et al., 2020). Originally, due to a combination of physio-chemical property assessments and a lack of observational data, OPEs were thought to have been a safe substitute. According to the Danish EPA in the EU Risk Assessment they found that TPP (an OPE) didn’t appear to have a greater negative impact than the other  flame retardants (FRs) they were using at the time (Pakalin et al., 2007). Other research has also used this same argument, citing that OPEs do not meet the criteria listed out by the Stockholm Convention to be as harmful as the now restricted/banned polybrominated diphenyl ethers (PBDEs) that they are replacing(Fiedler et al., 2019). Under the Stockholm convention there are 4 key criteria listed to be qualified as persistent organic pollutants (POPs) they are: 1) persistence potential i.e persists in the environment and is resistant to degradation, 2) bioaccumulation potential i.e each chemical’s potential to accumulate within various organisms, 3) their long-range transport potential (LRTP) i.e the chemical’s ability to propagate itself beyond local release locations and 4) Toxicity (Fiedler et al., 2019). The criterion is determined by both observational data and modelling tools. At the time of writing there are modelling tools that underestimate the POP like capabilities of OPEs (Sühring et al., 2020).  While on the observational side, it is becoming apparent that OPEs share POP-like characteristics; for example one study found that OPEs persist in indoor dust and bioaccumulate within humans (Y. Wang et al., 2021). 
Part of these issues in properly predicting POP like properties of OPEs may be attributed to the modelling tools used to assess OPEs. Default parameters of tools such as the Organization of Economic Co-operation and Development (OECD) Persistance (POV) and LRTP Screening Tool (“The Tool”) underestimated both the POV and LRTP values for OPEs(Sühring et al., 2020).  This is incredibly alarming as predictive modelling tools such as the OECD and POV Screening tool uses data such as OPEs having have half-lives of less than 2 days in air to predict LRTP. Simply based on the knowledge of OPE half-lives being less than 2 days in air, the sheer notion that these OPEs could be found in remote places such as the Arctic would be absurd; and yet emerging data clearly attests to this  (Sühring et al, 2021)(Sühring et al., 2016).  One way this disparity can be sufficiently explained then is through aerosol absorption. By absorbing into aerosols those same OPEs now had a mode of transport over long ranges (Möller et al., 2012). As noted by Möller et al the half lives presented is only within the gaseous state, this is a factor that must be analyzed more thoroughly to better understand the half-lives of OPEs not just in the gaseous state, but when adsorbed to other aerosols. 

# Chapter 2: OECD and POV Screening TOol

One commonly used assessment tool for LRTP and toxicity is the OECD POV and LRTP Screening Tool (The Tool). Designed with extensive research of PBDEs and other POPs in mind the Tool was parameterized to estimate the persistence and LRTP of POPs but would come short in adequately quantifying the persistence and LRTP of many more hydrophilic OPFRs (Zhang and Sühring et al., 2016). Multiple comparative models between The Tool and other models have begun to show the limitations of said parameters. One study found that the poly parameter linear free energy relationships (ppLFERs) tool had the ability to improve the current gas-particle partitioning description for polar chemicals(CW et al., 2008). Another study found that by including other environmental parameters such as wind speed had a significant effect on LRTP(Zhu et al., 2014). Clearly from this it can be seen that the tool must be updated, its ability to predict the fate of OPEs must be broadened as all parameters must be reassessed. This is the goal that this paper sought out attempted to solve as it can be seen throughout the rest of this paper. 


```{r include=FALSE}


#all packages, libraries and working directories used
setwd("C:/Users/monst/Desktop/master stuff/research stuff/R learning/ES8913/Final_project/TEST2/Final-Project-Rahmi-C")
list.files()
knitr::opts_chunk$set(echo = FALSE)
library(gridExtra)
library(ggplot2)
library(lubridate)
library(dplyr)
library(tidyverse)
library(tidyverse)
library(knitr)
library(GGally)
library(wesanderson)
library(scales)
install.packages("ggplot2")
install.packages("GGally")
install.packages ("dplyr")
install.packages ("tidyverse")
install.packages("wesanderson")
library(gridExtra)
library(ggplot2)
library(lubridate)
library(dplyr)
library(tidyverse)
library(knitr)
library(GGally)
library(wesanderson)
library(scales)



```

# Chapter 3 Figures and Results 

```{r message=FALSE, warning=FALSE, fig.width= 7,fig.height= 8}
#This data chunk is meant to begin everything, it loads the CSV file, proceeds to name it as a tible, changes the column names for a better easier visual look, then attempts to make a table using kable, i tried everything to make this table, but for some reason during the knit output the table always ended up looking a little smudged together, this was the best that I could get it at. Just in case every now and again, I include the libraries within the chunks. 


OPEdata<-read_csv("OPEdata.csv", col_names = TRUE) 
OPEdata <-as_tibble(OPEdata)
library(knitr)
OPE_ggcor <- OPEdata #it was named OPE_ggcor for the upcoming chart as it was the best way to read that chart
colnames(OPE_ggcor)[colnames(OPE_ggcor) %in% c("CTD_in_air", "CTD_in_water", "POV_air", "POV_water", "Carbon", "Hydrogen", "Chlorine", "Oxygen", "Phosphorus", "EPI_Suite_KAW","EPI_Suite_KOW","EPI_Suite_KOA")] <-c("CTD_a", "CTD_w", "POV_a","POV_w", "H","C","Cl", "O","Ph","KAW","KOW","KOA")
kable(OPE_ggcor, 
      caption = "Summary of OPE Data",
      row.names = FALSE, 
      longtable = TRUE, 
      align = c('l', 'c', 'c', 'c','c', 'c', 'c','c', 'c', 'c','c', 'c', 'c','c', 'c'), #attempted so many attempts at this
      booktabs = T, 
format.args = list(),
 escape = TRUE)
      
```
# 3.1 Table Results 
  As it can be seen from the table, there are a total of 14 variables being compared to here for the given 10 OPEs. Not all data was present for each OPE so some appear blank.


```{r message=FALSE, warning=FALSE, fig.width= 8,fig.height= 8}
# The second main code chunk, here ggcorr was used to make an easy to read Correlation Chart. All the variables were pitted against each other, oriented at an angle to overlap less, and label = T was used to show the values within the box. Labs used to label them and edit sizes.  

ggcorr(
  OPE_ggcor[, -1],
  method = c("pairwise","pearson"),
  max_size = 6,
  nbreaks = 10,
  size = 3,
  hjust = 0,
  angle = -45,
  layout.exp = 1,
  label = T,
) + ggplot2::labs(title = "Figure 1: OPE Correlation Chart", caption = "Correlation Data for all 13 variables analyzed for all OPEs")+
   theme(plot.caption= element_text(size=15)) +
  coord_flip(clip = "off")
  
```

# 3.2 Chart Results
An easy to read summary data chart of the correlation potential between the OPE variables. The chart starts with mass on the left, all the way down to RT on the right. The chart was broken down into 10 categories of correlation values, with +/-(0.8,1) having the highest degree of absolute coefficients. Due to rounding to 1 significant digit some values have a complete correlation. 


```{r message=FALSE, warning=FALSE, fig.width= 7,fig.height= 8, setup = FALSE}

#first use of ggpairs, the first column of OPEs was excluded. The method was pearson, the upper size was attempted to wrapped to 2.5 but this would not work, strange because it worked for the others. There was an attempt to separate the scale on the x-axis using scale_x_discrete yet sadly no matter how much I tried, it didn't spearte. Instead upon use as you can see it would simply make the scale disappear for unknown reasons

ggpairs(
OPE_ggcor[2:15],
method = c("pairwise","pearson"),
 max_size = 6,
  size = 3,
upper = list(continous = wrap("cor", size = 2.5)),
  hjust = -1,
  label = T,
) + ggplot2::labs(title = "Figure 2: OPE Correlation Chart of all Variables", caption = "Correlation Data for all 14 variables analyzed for all OPEs")+
   theme(plot.caption= element_text(size=15, hjust = 0)) +
  scale_x_discrete(guide = guide_axis(n.dodge=3))
        

```

# 3.3 GGPairs OPE Correlation Data. 
Here we can see the total analysis of all the OPEs pitted against each other. Due to the sheer number of variables it becomes difficult to read so this chart is simply used as a demonstration. More comprehensive graphs are designed below breaking down the variables within smaller categories 

  
 

```{r message=FALSE, warning=FALSE, fig.width= 7,fig.height= 8} 
# the first subset was made using columns to pick out the variables I wanted analyzed. Then upper and lower functions were used to determine, size and colour
OPE_mass <- ggpairs(OPE_ggcor, 
              package = "reshape",
              method = c("pairwise","pearson"),
              hjust = 0,
              upper = list(continuous = wrap("cor", size = 2.5)),
              columns = c("mass", "KAW", "KOW", "KOA","CTD_a", "CTD_w", "POV_a","POV_w"),
              lower = list(continuous = wrap("points", color = "blue", alpha = 0.5), 
                     combo = wrap("box", color = "blue", alpha = 0.3), 
                     discrete = wrap("facetbar", color = "yellow", alpha = 0.3) ), 
        diag = list(continuous = wrap("densityDiag",  color = "red", alpha = 0.5) )) +
  ggplot2::labs(title = " Subset# 1: OPE Correlation Chart with 8 Variables ", caption = "The 8 variables being anaylzed here for correlation values are: mass, \n \n KAW, KOW, KOA, CTD_w, CTD_a, POV_a, and POV_w")+
   theme(plot.caption= element_text(size=10, hjust = 0)) #+
  #scale_x_discrete(guide = guide_axis(n.dodge=3))
#scale_y_discrete(guide = guide_axis(n.dodge=3))
OPE_mass

```
# 3.4 GGPlots Subset Chart 1.
This subset compares KAW, KOW, KOA, CTD_w, CTD_a, POV_a,POV_w, and mass.  Some notable values here are between the pairs: CTD_w-POV_w, KOA-mass,KOA-KAW, and CTD_a-KOA with values: 0.999,0.910,-0.891 and -0.944 respectively. 

```{r message=FALSE, warning=FALSE, fig.width= 7,fig.height= 8} 
# the same parameters were applied to this subset, only wtih different columns being analyzed this time 
OPE_chemical <- ggpairs(OPE_ggcor, 
              package = "reshape", 
              method = c("pairwise","pearson"),
              hjust = 0,
              upper = list(continuous = wrap("cor", size = 2.5)),
              columns = c("RT","H","O","C","Cl", "KAW", "KOW", "KOA"),
              lower = list(continuous = wrap("points", color = "yellow", alpha = 0.5), 
                     combo = wrap("box", color = "green", alpha = 0.3), 
                     discrete = wrap("facetbar", color = "yellow", alpha = 0.3) ), 
        diag = list(continuous = wrap("densityDiag",  color = "purple", alpha = 0.5) )) +
 ggplot2::labs(title = " Subset# 2: OPE Correlation Chart with 8 Variables ", caption = "The 8 variables being anaylzed for correlation values are: RT, \n \n H, O, C, Cl, KAW, KOW, KOA")+
   theme(plot.caption= element_text(size=10, hjust = 0)) #+
  #scale_x_discrete(guide = guide_axis(n.dodge=3))+ #could not separte properly 
  #scale_y_discrete(guide = guide_axis(n.dodge=3))
OPE_chemical

```
# 3.5 GGPlots Subset Chart 2.
This subset compares "RT","H","O","C","Cl", "KAW", "KOW", and "KOA".  Some notable pairs here not already previously mentioned are: KOA-RT,KOA-H,KOA-C, and KOW-H, with values: 0.969,0.952,0.921, and 0.909 respectively. #note how significant this is because the pairs KOA-H and KOA-C may be a good predictive factor for what mode of transport other similar OPEs may use.  With OPEs high in Carbon, oxygen and hydrogen likely to be in air while not so much in water. While OPEs with Cl have a higher likely hood of using water as their mode of transport 

```{r message=FALSE, warning=FALSE, fig.width= 7,fig.height= 8} 
#once more only with a different set of variables, overlaps of results are expected.
OPE_chemical_res <- ggpairs(OPE_ggcor, 
              package = "reshape",  
              method = c("pairwise","pearson"),
              upper = list(continuous = wrap("cor", size = 2.5)),
              columns = c("RT","H","O","C","Cl", "CTD_a", "CTD_w", "POV_a","POV_w"),
              lower = list(continuous = wrap("points", color = "green", alpha = 0.5), 
                     combo = wrap("box", color = "orange", alpha = 0.3), 
                     discrete = wrap("facetbar", color = "green", alpha = 0.3) ), 
        diag = list(continuous = wrap("densityDiag",  color = "brown", alpha = 0.5) ))+
 ggplot2::labs(title = " Subset# 3: OPE Correlation Chart with 9 Variables ", caption = "The 9 variables being anaylzed for correlation values are: RT, \n \n H, O, C, Cl, CTD_a, CTD_W, POV_a, POV_w")+
   theme(plot.caption= element_text(size=10, hjust = 0)) #+
  #scale_x_discrete(guide = guide_axis(n.dodge=3))
#scale_y_discrete(guide = guide_axis(n.dodge=3))
OPE_chemical_res


```
# 3.6 GGPlots Subset Chart 3.
This subset compares RT","H","O","C","Cl", "CTD_a", "CTD_w", "POV_a","POV_w". The main point of this subset is to compare the chemical compositions with the POV.  Some notable pairs here not already previously mentioned are: POV_a-POV_w, and POV_a-O, with values: 0.999, 0.993 respectively. #note how significant this is, the POV_a match up with O at such a high value matches up with its CTD_a-O as well. Showing that O is an indicator of both and likely needed for atmospheric transport

```{r message=FALSE, warning=FALSE, fig.width= 7,fig.height= 8} 
#slight different, this subset broke up the OPEs by mass. Anything under 300 in category, anything over 300 in another. This was done using subset.This was also done to see whether correlation values would differ by much when isolated for smaller masses

OPE_lower_mass <-subset(OPE_ggcor, OPE_ggcor$mass <300)

OPE_lower_mass_chart <- ggpairs(OPE_lower_mass, 
              method = c("pairwise","pearson"),
              package = "reshape",  
              upper = list(continuous = wrap("cor", size = 2.5)),
              columns = c("mass","RT", "CTD_a", "CTD_w", "POV_a","POV_w","KAW","KOW"),
              lower = list(continuous = wrap("points", color = "green", alpha = 0.5), 
                     combo = wrap("box", color = "orange", alpha = 0.3), 
                     discrete = wrap("facetbar", color = "green", alpha = 0.3) ), 
        diag = list(continuous = wrap("densityDiag",  color = "brown", alpha = 0.5) ))+
 ggplot2::labs(title = " Subset# 4: OPE Lower Mass Correlation Chart with 9 Variables ", caption = "The 9 variables being anaylzed for correlation values are: RT, \n \n H, O, C, Cl, CTD_a, CTD_W, POV_a, POV_w")+
   theme(plot.caption= element_text(size=10, hjust = 0)) #+
  #scale_x_discrete(guide = guide_axis(n.dodge=3))
#scale_y_discrete(guide = guide_axis(n.dodge=3))
OPE_lower_mass_chart

```
# 3.7 GGPlots Subset Chart 4.
This subset compares "mass","RT", "CTD_a", "CTD_w", "POV_a","POV_w","KAW","KOW". Due to the absence in all data for each OPE, this data is less accurate for the representation of OPEs with a mass less than 300. Despite this,some notable pairs here  are: KOW-RT(0.952),KOW-CTD_w (-0.926), KOW-POV_w(-0.931), CTD_a-mass(-0.912),and CTD_a-RT(-0.941) with values: 0.952, -0.926,-0.931,-0.912,-0.941 respectively. 

```{r message=FALSE, warning=FALSE, fig.width= 7,fig.height= 8} 
# the last of the datasets looking at everything over 300
OPE_higher_mass <- subset(OPE_ggcor, OPE_ggcor$mass >300)

OPE_higher_mass_chart <- ggpairs(OPE_higher_mass, 
              method = c("pairwise","pearson"),
              package = "reshape",  
              upper = list(continuous = wrap("cor", size = 2.5)),
              columns = c("mass","RT", "CTD_a", "CTD_w", "POV_a","POV_w","KAW","KOW"),
              lower = list(continuous = wrap("points", color = "green", alpha = 0.5), 
                     combo = wrap("box", color = "orange", alpha = 0.3), 
                     discrete = wrap("facetbar", color = "green", alpha = 0.3) ), 
        diag = list(continuous = wrap("densityDiag",  color = "brown", alpha = 0.5) ))+
 ggplot2::labs(title = " Subset# 5: OPE Higher Mass Correlation Chart with 9 Variables ", caption = "The 9 variables being anaylzed for correlation values are: RT, \n \n H, O, C, Cl, CTD_a, CTD_W, POV_a, POV_w")+
   theme(plot.caption= element_text(size=10, hjust = 0)) #+
  #scale_x_discrete(guide = guide_axis(n.dodge=3))
#scale_y_discrete(guide = guide_axis(n.dodge=3))
OPE_higher_mass_chart

```
# 3.8 GGPlots Subset Chart 5.
This subset compares "mass","RT", "CTD_a", "CTD_w", "POV_a","POV_w","KAW","KOW". Due to the absence in all data for each OPE, this data is less accurate for the representation of OPEs with a mass over 300. Despite this,some notable pairs here  are: POV_a-KAW, and POV_w-CTD_w  with values: -0.997 and 0.998 respectively.

# Conclusion and Discussion

As noted throughout the entirety of the results, many of the variables within the OPE parameters have some sort of relationship with each other, whether that be positive or negative. Looking deeper into these results in section 3.5 note how significant the statistical relationship is between the pairs KOA-H and KOA-C. This is important because it may be a good predictive factor for what mode of transport other similar OPEs with high units of hydrogen and carbon atoms may use.  Implying that OPEs high in Carbon, oxygen and hydrogen likely to be in air while not so much in water. Section 3.5 also highlights to a lesser degree that OPEs with Cl have a higher likely hood of using water as their mode of transport.In section 3.6 there was also another pair with statistical significance: POV_a-O. This is because it supports the previous findings of 3.5, higher concentrations of oxygen have a strong positive coefficient value with the overall persistence of the OPE in air. Implying once more that this type of OPE is more likely to use an atmospheric mode of transport. This adds evidence to the fact that specific compounds may be indicators for certain modes of transports.These results are profound in that sense that when isolated for each significant relationship a crude model can be made to predict future OPEs. Future work I would love to do is to attempt these concepts on other OPEs to see how well it stands up, as well as expand the database to get more accurate results. 
 

 





## References
Castro-Jiménez, J., Berrojalbiz, N., Pizarro, M., & Dachs, J. (2014). Organophosphate ester (OPE) flame retardants and plasticizers in the open Mediterranean and Black Seas atmosphere. Environmental Science & Technology, 48(6), 3203–3209. https://doi.org/10.1021/ES405337G

Chokwe, T. B., Abafe, O. A., Mbelu, S. P., Okonkwo, J. O., & Sibali, L. L. (2020). A review of sources, fate, levels, toxicity, exposure and transformations of organophosphorus flame-retardants and plasticizers in the environment. Emerging Contaminants, 6, 345–366. https://doi.org/10.1016/J.EMCON.2020.08.004

Chupeau, Z., Bonvallot, N., Mercier, F., Bot, B. Le, Chevrier, C., & Glorennec, P. (2020). Organophosphorus Flame Retardants: A Global Review of Indoor Contamination and Human Exposure in Europe and Epidemiological Evidence. International Journal of Environmental Research and Public Health, 17(18), 1–24. https://doi.org/10.3390/IJERPH17186713

CW, G., M, S., M, M., F, W., U, S., & K, H. (2008). Dependence of persistence and long-range transport potential on gas-particle partitioning in multimedia models. Environmental Science & Technology, 42(10), 3690–3696. https://doi.org/10.1021/ES702619P

Du, J., Li, H., Xu, S., Zhou, Q., Jin, M., & Tang, J. (2019). A review of organophosphorus flame retardants (OPFRs): occurrence, bioaccumulation, toxicity, and organism exposure. Environmental Science and Pollution Research, 26(22), 22126–22136. https://doi.org/10.1007/S11356-019-05669-Y/FIGURES/2 

Fiedler, H., Kallenborn, R., Boer, J. de, & Sydnes, L. K. (2019). The Stockholm Convention: A Tool for the Global Regulation of Persistent Organic Pollutants. Chemistry International, 41(2), 4–11. https://doi.org/10.1515/CI-2019-0202

Guigueno, M. F., & Fernie, K. J. (2017). Birds and flame retardants: A review of the toxic effects on birds of historical and novel flame retardants. Environmental Research, 154, 398–424. https://doi.org/10.1016/J.ENVRES.2016.12.033

Lai, S., Xie, Z., Song, T., Tang, J., Zhang, Y., Mi, W., Peng, J., Zhao, Y., Zou, S., & Ebinghaus, R. (2015). Occurrence and dry deposition of organophosphate esters in atmospheric particles over the northern South China Sea. Chemosphere, 127, 195–200. https://doi.org/10.1016/J.CHEMOSPHERE.2015.02.015

Luo, P., Bao, L. J., Guo, Y., Li, S. M., & Zeng, E. Y. (2016). Size-dependent atmospheric deposition and inhalation exposure of particle-bound organophosphate flame retardants. Journal of Hazardous Materials, 301, 504–511. https://doi.org/10.1016/J.JHAZMAT.2015.09.014

Möller, A., Xie, Z., Caba, A., Sturm, R., & Ebinghaus, R. (2011). Organophosphorus flame retardants and plasticizers in the atmosphere of the North Sea. Environmental Pollution, 159(12), 3660–3665. https://doi.org/10.1016/J.ENVPOL.2011.07.022

Möller, A., Sturm, R., Xie, Z., Cai, M., He, J., & Ebinghaus, R. (2012). Organophosphorus Flame Retardants and Plasticizers in Airborne Particles over the Northern Pacific and Indian Ocean toward the Polar Regions: Evidence for Global Occurrence. Environmental Science and Technology, 46(6), 3127–3134. https://doi.org/10.1021/ES204272V

Pakalin, S., Cole, T., Steinkellner, J., Nicolas, R., Tissier, C., Munn, S., & Eisenreich, S. (2007). “REVIEW ON PRODUCTION PROCESSES OF DECABROMODIPHENYL ETHER (DECABDE) USED IN POLYMERIC APPLICATIONS IN ELECTRICAL AND ELECTRONIC EQUIPMENT, AND ASSESSMENT OF THE AVAILABILITY OF POTENTIAL ALTERNATIVES TO DECABDE.” http://www.jrc.cec.eu.int

Salamova, A., Ma, Y., Venier, M., & Hites, R. A. (2013). High Levels of Organophosphate Flame Retardants in the Great Lakes Atmosphere. Environmental Science and Technology Letters, 1(1), 8–14. https://doi.org/10.1021/EZ400034N/SUPPL_FILE/EZ400034N_SI_001.PDF

Sühring, R., Diamond, M. L., Scheringer, M., Wong, F., Pućko, M., Stern, G., Burt, A., Hung, H., Fellin, P., Li, H., & Jantunen, L. M. (2016). Organophosphate esters in Canadian Arctic air: Occurrence, levels and trends. Environmental Science and Technology, 50(14), 7409–7415. https://doi.org/10.1021/ACS.EST.6B00365/SUPPL_FILE/ES6B00365_SI_001.PDF

Sühring, R., Scheringer, M., Rodgers, T. F. M., Jantunen, L. M., & Diamond, M. L. (2020). Evaluation of the OECD POV and LRTP screening tool for estimating the long-range transport of organophosphate esters. Environmental Science: Processes & Impacts, 22(1), 207–216. https://doi.org/10.1039/C9EM00410F

Zhao, L., Zhang, Y., Deng, Y., Jian, K., Li, J., Ya, M., & Su, G. (2020). Traditional and emerging organophosphate esters (OPEs) in indoor dust of Nanjing, eastern China: Occurrence, human exposure, and risk assessment. Science of The Total Environment, 712, 136494. https://doi.org/10.1016/J.SCITOTENV.2020.136494

Zhang, X., Sühring, R., Serodio, D., Bonnell, M., Sundin, N., & Diamond, M. L. (2016). Novel flame retardants: Estimating the physical–chemical properties and environmental fate of 94 halogenated and organophosphate PBDE replacements. Chemosphere, 144, 2401–2407. https://doi.org/10.1016/J.CHEMOSPHERE.2015.11.017

Zhu, Y., OR, P., S, T., KC, J., & AJ, S. (2014). A new multimedia contaminant fate model for China: how important are environmental parameters in influencing chemical persistence and long-range transport potential? Environment International, 69, 18–27. https://doi.org/10.1016/J.ENVINT.2014.03.020
